---
title: "Opinion Formation on the Internet"
date: "`r Sys.Date()`"
bibliography: ["rpackages.bib", "refs.bib"]
biblio-style: "apalike"
link-citations: true
output:
  rmdformats::material:
    highlight: kate
    collapsed: false
    dev: png
    self_contained: true
    number_sections: true
    code_folding: hide
    thumbnails: true
    gallery: true
    fig_width: 6
    fig_height: 3.5
    df_print: kable
---


```{r setup, cache=FALSE, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE, 
               fig.height = 5,
               fig.width = 7)
opts_knit$set(width=75)

library(tidyverse)
library(scales)
library(gt)
```



# Abstract
Today the majority of people uses online social networks not only to stay in contact with friends, but also find information about relevant topics, or to spread information. While a lot of research has been conducted into opinion formation, only little is known about what influences the spread of information in online social networks. To answer this question, we created an agent-based model and simulated message spreading in social networks using a latent-process model. In our model, we varied four different content types, five different network types, and we varied between a model that also includes a personality model of the agents and a model that did not. We found out, that the network type has only a weak influence on the distribution of content, where the message type and the personality of the user have a clear influence on how many users receive a message.  

```{r test-main, child = '10_intro.rmd'}
```


# Analysis

Using the agent-based model, we analyzed whether different initial
settings lead to different outcomes. As initial configurations, we used
different content types (*affective content*, *cognitive content*,
*both*, *weak content*) and different network types (*Facebook*,
*Barábasi-Albert*, *Watts Strogatz*, *Scale Free Network*, *Random
Network*). Further, we compared whether the use of a *personality model*
for the creation of the agents in our simulation leads to a different
outcome than the simulation runs without the *personality model*. Using
these different initial settings, we found the following results.

Our results are depicted in the figure below. It
shows for each simulation step how many of the agents who saw the
message also forwarded it. The number of forwarding agents is also
visible for the five different network types (horizontal); for the
agents with and without the use of the *personality model* (vertical);
and the different content types (color).

## Highest and lowest propoertion of forwarding agents

As can be seen in this figure, never all agents have seen and forwarded
the message. This applies to all initial settings. The number of
forwarding agents was highest in the simulation, where the content is
*both*, affective and cognitive, where the agents have a *personality*,
and where the agents are in the *Watts Strogatz* network. Using this
initial settings, more than 75% of the agents did forward the seen
content.

In contrast, the lowest number of forwarding agents occurred in the
simulation, where the content is *weak* or *mostly cognitive*, where the
agents are designed according to the *personality model* and where the
agents are located in a *Barábasi Albert* network. The agents stop
forwarding the message at the latest at the fourth simulation step and
until then almost no agent has forwarded the message.

So far we considered, when the proportion of forwarding agents is
highest or lowest. Following, we look at the single factors that could
have an influence on the proportion of forwarding agents, starting with
the content type.

```{r read_data, cache=TRUE}
filename <- "../output/results.csv"
results <- read_csv(filename) %>% 
  mutate(agent_generator = factor(agent_generator, 
                                  levels = c("generatePersonalityAgent", "generateRandomAgent"),
                                  labels = c("Personality", "Random")
                                  ),
         network_generator = factor(network_generator,
                                    levels = c("generateBarabasi", "generateFacebook", "generateRandomNetwork", "generateStochasticBlockModel", "generateScaleFree", "generateWattsStrogatz"),
                                    labels = c("Barábasi", "Facebook", "Random", "SBM", 
                                               "Cho",  "Watts Strogatz")),
         value = paste0(affective_value,"/",cognitive_value) 
         ) %>% mutate(content_type = factor(value, levels = rev(c("0.2/0.2", "0.2/0.8", "0.8/0.2", "0.8/0.8")),
                                     labels = rev(c("weak content", "mostly cognitive", "mostly affective", "both"))))

results$network_generator %>% factor() %>%  levels()
full_design <- expand.grid(step = unique(results$step),
                           agent_count = unique(results$agent_count),
            network_generator = unique(results$network_generator),
            runid = unique(results$runid),
            agent_generator = unique(results$agent_generator))

results <- full_design %>% left_join(results) %>% fill(everything(), start_agent, new_agents)
```


```{r get_degree_distribution, cache=TRUE}
plot_degree <- function(x) {
  as.numeric(str_split(x, pattern = "; ", simplify = TRUE))
}

x <- results[1,]$degree_centrality
plot(density(plot_degree(x)))

# clustering 
results %>% filter(step == 1) %>% 
  ggplot() + 
  aes(x = factor(agent_count), y = clustering_coefficient) + 
  geom_violin() + 
  geom_boxplot() +
  coord_flip() + 
  scale_y_log10() +
  #geom_point() + 
  facet_wrap(~network_generator) + 
  labs(x = "Network Size", y = "Clustering Coefficient (log)") +
  ggtitle("Distribution of Clustering Coefficient by Network Generator") +
  theme_bw(base_size = 14) -> p

ggsave("../output/clustering.pdf", p, width = 7, height = 5, dpi = 600)

# clustering 
results %>% filter(step == 1) %>% 
  ggplot() + 
  aes(x = factor(agent_count), y = community_count) + 
  geom_violin() +
  geom_boxplot() + 
  geom_point() +
  coord_flip() +
  scale_y_log10() +
  #geom_point() + 
  facet_wrap(~network_generator) +
  labs(x = "Network Size", y = "Community Count Coefficient (log)") +
  ggtitle("Distribution of Community Count by Network Generator") +
  theme_bw(base_size = 14) -> p

ggsave("../output/community.pdf", p, width = 7, height = 5, dpi = 600)

# Only the last step
```


```{r visualize_1, cache=TRUE}

count_select <- 3500
simulation_count <- results %>% 
  filter(agent_count > count_select, step == 1) %>% count() %>% pull()

max_step <- results %>% pull(step) %>% max()

(p <- results %>% 
  filter(agent_count > count_select) %>% 
  #sample_n(1000) %>% 
  ggplot() +
  aes(x = step, y = (seen+sent)/agent_count, group = interaction(runid, network_generator), color = content_type)+ 
    geom_line(alpha = 0.1) + 
  facet_grid(factor(agent_generator)~network_generator) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_colour_brewer(palette = "Set1") +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  labs(title = "Spread of messages by content type, agent, and network",
       subtitle = "Network penetration depends mostly on agent type and content",
       caption = paste0("Each line represents the spread of a message in one simulation (of ",simulation_count," simulations)"),
       x = "Simulation step", y = paste0("Proportion of forwarding agents (n = ",count_select,")"), color = "Content type") + 
    theme_bw())

ggsave("../output/spread1.png", p, width = 7, height = 5, dpi = 600)
```

```{r eval=FALSE, include=FALSE}
results %>% filter(runid == 1, step == 1)

full_design <- expand.grid(step = unique(results$step),
                           agent_count = unique(results$agent_count),
            network_generator = unique(results$network_generator),
            runid = unique(results$runid),
            agent_generator = unique(results$agent_generator))

test <- full_design %>% left_join(results) %>% fill(everything(), start_agent, new_agents)

```




```{r}
results %>% 
  filter(agent_count > count_select) %>% 
   group_by(network_generator, agent_generator, step, content_type) %>% 
   summarize(seen_p = mean((seen + sent)/agent_count),
             seen_sd = sd((seen + sent)/agent_count),
             simulations = n()) %>% 
  mutate(seen_se = seen_sd / sqrt(simulations) ) %>% 
  ggplot() +
  aes(x = step, y = seen_p, ymin = seen_p - seen_se, ymax = seen_p + seen_se, color = content_type) +
  geom_ribbon(alpha = 0.7, fill = "lightgray", size = 0.1) +
  geom_errorbar(alpha = 0.7) +
  geom_line(size = 0.1) +
  geom_point(size = 0.3, alpha = 0.7) +
  facet_grid(agent_generator~network_generator) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_colour_brewer(palette = "Set1") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)),
         ribbon = FALSE) +
  labs(title = "Spread of messages by content type, agent, and network",
       subtitle = "Network penetration depends mostly on agent type and content",
       caption = paste0("Ribbon and error bars denotes the standard error of the mean (of ",simulation_count," simulations)"),
       x = "Simulation step", y = paste0("Mean proportion of forwarding agents (n = ",count_select,")"), color = "Content type") + 
    theme_bw() -> p

p
ggsave("../output/spreadMSE.png", p, width = 7, height = 5, dpi = 600)
```


```{r visualize_2, cache=TRUE}

count_select <- 2000
simulation_count <- results %>% 
  filter(agent_count == count_select, step == 1) %>% count() %>% pull()

(p <- results %>% 
  filter(agent_count == count_select) %>% 
  #sample_n(1000) %>% 
  ggplot() +
  aes(x = step, y = (seen + sent)/agent_count, group = runid, color = content_type) + geom_line(alpha = 0.1) + 
  facet_grid(factor(agent_generator)~network_generator) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_colour_brewer(palette = "Set1") +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  labs(title = "Spread of messages by content type, agent, and network",
       subtitle = "Network penetration depends mostly on agent type and content",
       caption = paste0("Each line represents the spread of a message in one simulation (of ",simulation_count," simulations)"),
       x = "Simulation step", y = paste0("Proportion of forwarding agents (n = ",count_select,")"), color = "Content type") + 
    theme_bw())
p

ggsave("../output/spread2.png", p, width = 7, height = 5, dpi = 600)
```

```{r visualize_3, cache=TRUE}

count_select <- 1000
simulation_count <- results %>% 
  filter(agent_count == count_select, step == 1) %>% count() %>% pull()

(p <- results %>% 
  filter(agent_count == count_select) %>% 
  #sample_n(1000) %>% 
  ggplot() +
  aes(x = step, y = (seen + sent)/agent_count, group = runid, color = content_type) +
  geom_line(alpha = 0.1) + 
  facet_grid(factor(agent_generator)~network_generator) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_colour_brewer(palette = "Set1") +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  labs(title = "Spread of messages by content type, agent, and network",
       subtitle = "Network penetration depends mostly on agent type and content",
       caption = paste0("Each line represents the spread of a message in one simulation (of ",simulation_count," simulations)"),
       x = "Simulation step", y = paste0("Proportion of forwarding agents (n = ",count_select,")"), color = "Content type") + 
    theme_bw())

ggsave("../output/spread3.png", p, width = 7, height = 5, dpi = 600)
```


```{r, eval=FALSE}
results %>% 
  filter(step == 1) %>%  
  group_by(agent_generator, network_generator, agent_count)  %>% 
  summarize(mean_edges = mean(edge_count), 
            edges_se = sd(edge_count)/sqrt(n()),
            simulations = n()) %>% 
  ungroup() %>% 
  mutate(edges = paste0(format(mean_edges, digits = 1, big.mark = ",")," ±",format(edges_se, digits = 1))) %>% 
  select(-mean_edges, -edges_se) %>% 
  spread(agent_count, value = edges) %>% 
  gt() %>% 
  tab_header(paste0("Overview of network properties by configuration"),
             subtitle = paste0("In total ", format(simulation_count, big.mark = " "), " simulations were run using 10 different settings.")) %>%
   tab_row_group(
    group = "Personality based agents",
    rows = 1:6
    ) %>% 
  tab_row_group(
    group = "Random agents",
    rows = 7:12
    ) %>% 
  tab_spanner("Edges by network size (and standard error)", columns = vars("1000","2000","4039")) %>% 
  cols_hide(vars("agent_generator")) %>% 
  cols_label(network_generator = "Network type") %>% 
  cols_label("1000" = "1000 agents") %>% 
  cols_label("2000" = "2000 agents") %>% 
  cols_label("4039" = "4039 agents") %>% 
  cols_label("simulations" = "# of simulations") %>% 
  cols_align("right", columns = c("1000", "2000", "4039")) 
  
```


```{r read_data_3}
final_results <- results %>% group_by(runid, agent_generator, network_generator) %>% arrange(-step) %>% slice(1) 


final_results %>% 
  ggplot() +
  aes(x = network_generator, y = seen) + geom_boxplot()
```

## Content types
Comparing the four different contents, most agents forward the content,
that is *both* affective and cognitive. Agents forward the *mostly
affective* content the second most and significantly more frequently
than the other two contents. The *weak content* as well as the *mostly
cognitive content* are almost never forwarded at all. The agents also
always stop forwarding the message before the eighth simulation step.
Only with the *Facebook network* and without using the *personality
model*, the *mostly cognitive content* is forwarded somewhat more
frequently, but still forwarding does not exceed the eighth step of the
simulation.

## Network type
After considering the influence of the content type we now look at the
different network types and how they influence the number of forwarding
agents. As can be seen in the figure of the simulation, the
proportion of forwarding agents differs only slightly between the
*Random*, *Scale-Free*, and *Watts Strogatz* network. Hardly any
difference can be seen between the red and blue lines of the forwarding
agents located either in the *Random* or in the *Scale-Free* network.
The lines also look very similar for the agents in the *Watts Strogatz*
network. While in the *Random* and in the *Scale-Free* network many
agents already forward the *both content*, the number of forwarding
agents for the *Watts Strogatz* network is a bit higher. If we look at
the *mostly affective content*, the number of forwarding agents in the
*Watts Strogatz* network differs more for the individual simulations
than for the two previously mentioned networks.

Slightly larger differences can be seen for the *Barábasi Albert* and
the *Facebook* network. In the *Facebook* and the *Barábasi Albert*
network, the number of agents that forward the *both content* and the
*mostly affective content* is more similar. In the *Barábasi Albert*
network, compared to all other network types, fewer agents forward the
two most forwarded contents.

## Personality model
Lastly, we compare the message spread in our simulations based on
whether the personality of our agents followed a *personality model* or
was *randomly* generated.The figure of the simulation shows that the
proportion of forwarding agents was always higher when they were
equipped with a *personality model* in the *simulation*. The biggest
deviation occurs when the agents are located in the *Barábasi Albert
Network*.

Content, network and personality
--------------------------------

Overall, the figure of the simulation shows that the proportion of
forwarding agents mostly depends on the *content type* and if the agents
have a *personality* designed according to the *personality model* or
not. In contrast, the *network type* showed a lower influence except for
*Barábasi Albert* and *Facebook* networks.

## Correlation of affective and sending individuals
```{r cor}
cor(final_results$affective_value, final_results$sent)
```


## Linear model of parameters that play a role
```{r linear model}
lm(sent ~ affective_value + cognitive_value + agent_count + agent_generator + network_generator, data = final_results )
```


# Used Packages
The following R packages were used:

* Package: knitr	 by @R-knitr
* Package: tidyverse	 by @R-tidyverse
* Package: rmdformats	 by @R-rmdformats
* Package: DT	 by @R-DT
* Package: kableExtra	 by @R-kableExtra
* Package: scales	 by @R-scales
* Package: psych	 by @R-psych

`r if (knitr::is_html_output()) '# References {-}'`
